```{r}
#REMOVE This chunk and ensure that the line with the three dashes from below --- is on line 1 of this .rmd file
```
---
title: "Setting up .Rmd files"
author: "Lacey Hartigan"
date: "5/9/21"
output:
  pdf_document: default
  html_document: default
---

Note: If I want to knit to html instead of pdf, I move line 7 in the header above to line 6. Whichever one is first, that's the format .Rmd will knit to. 

```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#I tend to load all the libraries I will need for the whole .Rmd file first. Also, remember that whatever you load last will override the others IF there are overlapping commands. So, I tend to load tidyverse last. If you're getting weird issues - you can try loading packages in a different order
library(forcats)
library(knitr)
library(tidyverse)
```

Here's some great info on dealing with conflicting packages (and how to figure out if you have conflicts). [Dealing with Package Conflicts](https://www.tidyverse.org/blog/2018/06/conflicted/)

