---
title: "Regression, Specifying the Reference Group"
output: html_document
---

## Load needed libraries
```{r}
rm(list=ls())

library(tidyverse)
library(tidymodels)
```

## Add the "Area" dataset
```{r}
ad<-readRDS("area_data.Rds")
```

## Split the data into training and testing versions. 
```{r}
set.seed(524)

split_data<-ad%>%
  initial_split(prop=.5)

ad_train<-training(split_data)
ad_test<-testing(split_data)
```

## Set Linear Model

```{r}
lm_fit <- 
  linear_reg() %>% 
  set_engine("lm")%>%
  set_mode("regression")
```

## Begin Workflow

```{r}
income_wf<-workflow()%>%
  add_model(lm_fit)
```

## Set Formula
```{r}
income_formula<-as.formula("income_75 ~ college_educ + perc_homeown + metro + region")
```

## Set Recipe

Below, `step_dummy` is applied to both the `metro` and `region` variables (categorical variables). We can use step_relevel if we want to specify the reference group (i.e., the group that will be omitted from our results table). If you don't specify, R will choose the first group alphabetically or the 0/lowest-numbered group (if stored as numbers). 
```{r}
income_rec<-recipe(income_formula, data=ad_train)%>%
  step_relevel(region, ref_level = "South")%>% #here's where I tell R that I want South to be my reference
  step_dummy(metro, region)
```

## Update workflow with new recipe

Because we've already created the workflow `income_wf` we can update it with our new recipe, using the command `update_recipe`.
```{r}
income_wf<-income_wf%>%
  update_recipe(income_rec)
```

## Fit to training data and look at coefficients and model fit

Now we're ready to fit our linear model. The fit command below tells it to fit the model, with results stored in the object `lm_results`. We can then pipe the lm_results to the `tidy` command to see the coefficients. To see measures of model fit we can use `pull_workflow_fit` and then pipe those results to the `glance` command. 

```{r}
lm_results<-fit(income_wf, ad_train)

lm_results%>%
  tidy()

lm_results%>%
  extract_fit_parsnip()%>%
  glance()
```

## Using `last_fit` to pull RMSE and Rsquared

The `tidymodels` package has a function that automates the steps of running the model, generating predictions in the testing dataset and then generating metrics of model fit from the testing dataset. It's called `last_fit`.

```{r}
lf<-last_fit(income_wf, split=split_data)
lf$.metrics
```
