---
title: "Plots for Classifiers - LMS Module 9"
output:
  pdf_document: default
---

Plots are rarely used in the context of classification, but they can aid understanding. I'll show three ways of thinking about plots for classification: bar graphs, heatmaps, and plotting the probability predictions from a logit model.

We'll continue working with the random acts of pizza dataset. 
```{r}
library(tidyverse)
library(modelr)
library(knitr)

load("za.RData")
```

We always want to start with a cross tab of our dependent variable as a function of other variables. We structure cross tabs in a very particular way for the purposes of research: the independent variable goes on the rows, and the dependent variable goes on the columns. If proportions or percentages are going to be calculated, they should be calculated across rows. 

Recalling our previous lesson, let's look at a crosstab of `got_pizza` with the independent variable of `student`

```{r}
tab_student<-with(za, table(student, got_pizza)) #IV on rows, first; DV on columns, second
tab_student
```

If we want to make this a little better, we can change the row and column titles

```{r}
colnames(tab_student)<-c("No Pizza","Received a Pizza")
kable(tab_student) #kable makes a prettier table for R Markdown knit files
```

If we want to add proportions to this table, we can it like so:
```{r}
tab_student_prop<-prop.table(tab_student, margin=1) #margin=1 tells the command to make the rows sum to 1; if you did margin=2 it would sum by columns
kable(tab_student_prop) #pretties the table up for the final knit doc
```

Sometimes (okay, all the times) audiences prefer percentages. Easy enough to do:
```{r}
kable(round(tab_student_prop*100, 1)) #this rounds this to the tenths place because of the ,1 at the end (1 place after the decimal); 2 would make it to the hundredths place, etc. etc. 
```

If you want to include a third variable in a cross tab, that requires splitting the dataset. For instance, if we want to know the proportion of posts that include "student" AND "grateful" that received pizza, we would do this:
```{r}
#makes table filtered for ONLY those who had grateful in their post
tab_student_grateful<-with(filter(za,
                           as.character(grateful)=="Grateful in post"),
                           table(student, got_pizza))
tab_student_grateful
```

# Set up the code for the people who did NOT include grateful in post
```{r}
tab_student_notgrateful<-with(filter(za,
                                     as.character(grateful)=="Grateful not in post"), 
                              table(student, got_pizza))
tab_student_notgrateful
```

*Outcome by "Student" AND "Grateful"*
```{r}
prop.table(tab_student_grateful, margin=1)%>%
  kable()
#reminder, these are ONLY people who had "grateful" in their post
```


## Bar Graphs from Cross Tabs

To format the data for barcharts, we make use of the concept of conditional means. Let's use two variables to calculate the probability of receiving a pizza: poor and student. 

```{r}
za_sum<-za%>%
  group_by(poor, student)%>%
  dplyr::summarize(prob_pizza=mean(got_pizza, na.rm=TRUE))%>%
  ungroup()
za_sum
```

Then we can plot this using our familiar ggplot commands:
```{r}
#simple plot
gg1<-ggplot(za_sum, aes(y=prob_pizza, x=student, fill=poor))+
  geom_bar(stat="identity", position="dodge")
gg1

#prettying it up!
gg1plus<-ggplot(za_sum, aes(y=prob_pizza, x=student, fill=poor))+
  geom_bar(stat="identity", position="dodge")+ 
  xlab("Student Status") + ylab("Probability of Receiving a Pizza")+ 
  theme(legend.title=element_blank())+#removes title from legend 
  scale_fill_manual(values=c("Poor not in post"="light blue", "Poor in post"="dark green"))+ 
  geom_text(aes(label=round(prob_pizza, 2)), #add data labels to bars
                   position=position_dodge(width=.9), #matches positioning we originally specified above (position="dodge"); change the width# to adjust
                   vjust=-.25)+ #adjusts the height of the labels relative to the bars
  theme_classic()+ #removes the gray background and grid/etc. 
    scale_y_continuous(expand=c(0,0), limits=c(0, max(za_sum$prob_pizza)+.01)) #do you hate the "floating" bars like I do? If so, this line will stop the floating madness (the expand part) and it also ensures that you can see all data labels (the limits part)
gg1plus
```

## Heat Maps

To generate a heat map, we'll first divide up the independent variables into quintiles:

```{r}
#score = sentiment score (higher is more positive); karma = poster's karma on reddit
za<-za%>%
  mutate(score_quintile=ntile(score, 5), #5 is quintile
         karma_quintile=ntile(karma, 5))
#look at the data after you add these quintile vars - use the table function to make sure it worked properly
```

Then we'll create a summary dataset that shows the probabilities of the outcome across all of the combined categories of the two independent variables. 
```{R}
za_sum<-za%>%
  group_by(score_quintile, karma_quintile)%>% #group by two IVs
  dplyr::summarize(prob_pizza=mean(got_pizza, na.rm=TRUE))%>% #take mean of DV by this grouping
  arrange(-prob_pizza)%>% #arrange data by pred.probabilities (descending order)
  ungroup() 
za_sum

#notice that we have some data that are missing on our score/karma quintiles (this is because they were missing on the original variables). I don't want them in my final table/displays/etc. so I'm going to modify the code above to drop them
za_sum2<-za%>%
  group_by(score_quintile, karma_quintile)%>% 
  drop_na()%>% #this is going to drop anything that's missing on my two group_by variables
  dplyr::summarize(prob_pizza=mean(got_pizza, na.rm=TRUE))%>% 
  arrange(-prob_pizza)%>% 
  ungroup() 
za_sum2
```

Another way we could have dropped missing data is using filter 
```{r}
za_sum<-za_sum%>%
  filter(!(is.na(score_quintile)),!(is.na(karma_quintile))) #!is.na = not missing
za_sum
```

Now we're ready to plot!

```{r}
gg<-ggplot(za_sum,
           aes(x=as.factor(score_quintile),
               y=as.factor(karma_quintile), fill=prob_pizza))
gg<-gg + geom_tile() #heat map grid (tiles)
gg<-gg + scale_fill_gradient(low="white", high="darkgreen") #fill with gradient coloring where lowest probabilities are white and my high are dark green
gg<-gg + xlab("Emotional Valence") + ylab("Reddit Karma") #add labels
gg<-gg + theme(legend.title=element_blank()) #remove legend title

#we removed the legend title, but that means we don't tell the reader what is actually being plotted with our colors; add a title to do this
gg<-gg + ggtitle("Predicted Probability of Receiving a Pizza") + 
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=.5))

gg
```


## Plotting by probabilities from models

It can be difficult to plot the results of a logistic regression. We're going to use the same solution that we used for linear regression, where we create simulations from a hypothetical dataset. 

First we rerun our logistic regression. 
```{r}
logit_mod<-glm(got_pizza ~ karma + total_posts + raop_posts + student + grateful,
               data=za,
               na.action=na.exclude,
               family=binomial(link="logit"),
               y=TRUE)

#GENERAL CODE:
#logit_mod<-glm(y~x1+x2,
#               family=binomial(link="logit"))

summary(logit_mod)
```

Then we create some hypothetical data. 

```{r}
#create a dataset that would look like we want it to (we're constraining the data); then run our predictive model in it.
#going to specify a variable for each IV in the model
hypo_data<-data_grid(za, 
                     total_posts=seq_range(total_posts, n=100), #go from min to max of total posts in 100 steps
                     karma=mean(karma, na.rm=TRUE), #hold constant @ the mean of karma
                     raop_posts=mean(raop_posts, na.rm=TRUE), #hold constant @ mean # posts
                     grateful=levels(grateful)[1], #hold constant at first level of grateful; = did not indicate grateful
                     student=levels(student))%>% #lets student vary 
  mutate(pred=predict(logit_mod, newdata=., type="response")) #generate a prediction from our logit model

#we're letting total # posts var and the student variable vary (0 or 1)
```

And now we're ready to plot. 
```{r}
gg_student_ttlposts<-ggplot(hypo_data,
           aes(x=total_posts, y=pred, color=student))+ 
  geom_line()+ 
  xlab("Total Posts") + ylab("Predicted Probability of Pizza")+ 
  scale_color_manual(values=c("No student"="magenta", "Student"="black"))+ 
  theme(legend.title=element_blank()) #remove legend title - not needed in this case

gg_student_ttlposts
```

## How to do the same with Random Acts of Pizza posts (raop) varying
```{r}
#we're going to let RAOP and student vary, but nothing else
hypo_data2<-data_grid(za, 
                     total_posts=mean(total_posts, na.rm=TRUE), #hold ttl posts constant
                     karma=mean(karma, na.rm=TRUE), #hold karma constant
                     raop_posts=seq_range(raop_posts, n=100), #let RAOP vary
                     grateful=levels(grateful)[1], #hold grateful constant
                     student=levels(student))%>% #let student vary
  mutate(pred=predict(logit_mod, newdata=., type="response"))
```


```{r}
gg_student_raopposts<-ggplot(hypo_data2,
           aes(x=raop_posts, y=pred, color=student))+ 
  geom_line()+ 
  xlab("RAOP Posts") + ylab("Predicted Probability of Pizza")+ 
  scale_color_manual(values=c("No student"="cyan", "Student"="black"))+ 
  theme(legend.title=element_blank())

gg_student_raopposts

#This is the plot of predicted probability of Pizza Received by RAOP posts and whether poster was student or not (holding ttl posts, karma, and grateful constant).
```