---
title: "Assignment 4 - LMS Module 7"
author: "YOUR NAME HERE"
date: "due 10/27/2020"
output: pdf_document
---

1. Download data for all of the counties in Tennessee (TN) on highest education received.
```{r}

```

2. Compute the proportion of the population that has an associate’s degree or above in each TN county. Print the first several rows of your new dataset (hint: use the head() function). Do NOT print the whole dataset, please.
```{r}

```

3. Download data for all of the counties in TN on household income.
```{r}

```


4. Compute the proportion of the population that has a household income above $50,000 in each TN county. Print the first several rows of your new dataset (hint: use the head() function). Do NOT print the whole dataset, please.
```{r}

```


5. Create a scatterplot of the proportion of the population that has at least an associate's degree (x-axis) and  the proportion of the population with an income over $50,000 (y-axis).
```{r}

```

 
6. Create a map of the data you calculated in either problem 2 or 4 (education or income) (or create maps for both if you really want to flex those coding muscles!). 
```{r}

```
Provide a few sentences describing some interesting or unexpected findings you see from your map(s). 




